fGFPCA dynamic prediction

### 2023-02-15 f-GFPCA.Rmd
more updates interval prediction
model evaluation metrics

```{r, results='hide', message=FALSE}
source(here("Code/OutSampMLE.R"))
source(here("Code/OutSampBayes.R"))
```

@@

```{r, laplace_ped}
# clean MLE out-of-sample prediction results
out_pred_clean_laplace <- function(subjID=1, mid=mid){
  # Prediction
  pred_t195 <- out_pred_laplace(fpca_fit = fpca_fit, df_new = df %>% filter(id==subjID & sind_inx<=195) %>% select(-eta_i))
  pred_t395 <- out_pred_laplace(fpca_fit = fpca_fit, df_new = df %>% filter(id==subjID & sind_inx<=395) %>% select(-eta_i))
  pred_t595 <- out_pred_laplace(fpca_fit = fpca_fit, df_new = df %>% filter(id==subjID & sind_inx<=595) %>% select(-eta_i))

@@

```{r, results='hide', message=FALSE}
out_pred_lap_subj1 <- out_pred_clean_laplace(1, mid) %>% 
  left_join(df_true_eta %>% filter(id==1), by = "mid")
out_pred_lap_subj2 <- out_pred_clean_laplace(2, mid) %>% 
  left_join(df_true_eta %>% filter(id==2), by = "mid")
out_pred_lap_subj3 <- out_pred_clean_laplace(3, mid) %>% 
  left_join(df_true_eta %>% filter(id==3), by = "mid")
out_pred_lap_subj4 <- out_pred_clean_laplace(4, mid) %>% 
  left_join(df_true_eta %>% filter(id==4), by = "mid")
```

```{r, fig.width=12, fig.height=4}
# compare point prediction with in-sample
# subject 1
grid.arrange(
  out_pred_subj1 %>% 
    group_by("type") %>%
    ggplot()+
    geom_line(aes(x=mid, y=true), col = "black")+
    geom_line(aes(x=mid, y=eta, col= type, group = type), linetype="dashed", na.rm = T,
              show.legend = FALSE)+
    labs(x="Binned time", title = "Subject 1 out-of-sample MLE"),

  out_pred_lap_subj1 %>% 
    group_by("type") %>%
    ggplot()+
    geom_line(aes(x=mid, y=true), col = "black")+
    geom_line(aes(x=mid, y=eta, col= type, group = type), linetype="dashed", na.rm = T,
              show.legend = FALSE)+
    labs(x="Binned time", title = "Subject 1 out-of-sample Laplace"),

@@

  out_pred_lap_subj2 %>% 
    group_by("type") %>%
    ggplot()+
    geom_line(aes(x=mid, y=true), col = "black")+
    geom_line(aes(x=mid, y=eta, col= type, group = type), linetype="dashed", na.rm = T,
              show.legend = FALSE)+
    labs(x="Binned time", title = "Subject 2 out-of-sample Laplace"),

@@

  out_pred_lap_subj3 %>% 
    group_by("type") %>%
    ggplot()+
    geom_line(aes(x=mid, y=true), col = "black")+
    geom_line(aes(x=mid, y=eta, col= type, group = type), linetype="dashed", na.rm = T,
              show.legend = FALSE)+
    labs(x="Binned time", title = "Subject 3 out-of-sample Laplace"),

@@

  out_pred_lap_subj4 %>% 
    group_by("type") %>%
    ggplot()+
    geom_line(aes(x=mid, y=true), col = "black")+
    geom_line(aes(x=mid, y=eta, col= type, group = type), linetype="dashed", na.rm = T,
              show.legend = FALSE)+
    labs(x="Binned time", title = "Subject 4 out-of-sample Laplace"),

@@

@@

>>>>

```{r, results='hide', message=FALSE}
source(here("Code/OutSampMLE.R"))
# source(here("Code/OutSampBayes.R"))
```

#### Figure

@@

```{r, laplace_ped, eval=FALSE}
# clean MLE out-of-sample prediction results
out_pred_clean_laplace <- function(subjID=1, mid=mid){
  # Prediction
  pred_t195 <- out_pred_laplace(fpca_fit = fpca_fit, df_new = df %>% filter(id==subjID & sind_inx<=195) %>% select(-eta_i))
  pred_t395 <- out_pred_laplace(fpca_fit = fpca_fit, df_new = df %>% filter(id==subjID & sind_inx<=395) %>% select(-eta_i))
  pred_t595 <- out_pred_laplace(fpca_fit = fpca_fit, df_new = df %>% filter(id==subjID & sind_inx<=595) %>% select(-eta_i))

@@

```{r, results='hide', message=FALSE, eval=FALSE}
out_pred_lap_subj1 <- out_pred_clean_laplace(1, mid) %>% 
  left_join(df_true_eta %>% filter(id==1), by = "mid")
out_pred_lap_subj2 <- out_pred_clean_laplace(2, mid) %>% 
  left_join(df_true_eta %>% filter(id==2), by = "mid")
out_pred_lap_subj3 <- out_pred_clean_laplace(3, mid) %>% 
  left_join(df_true_eta %>% filter(id==3), by = "mid")
out_pred_lap_subj4 <- out_pred_clean_laplace(4, mid) %>% 
  left_join(df_true_eta %>% filter(id==4), by = "mid")
```

```{r, fig.width=8, fig.height=4}
# compare point prediction with in-sample
# subject 1
grid.arrange(
  out_pred_subj1 %>% 
    group_by("type") %>%
    ggplot()+
    geom_line(aes(x=mid, y=true), col = "black")+
    geom_line(aes(x=mid, y=eta, col= type, group = type), linetype="dashed", na.rm = T,
              show.legend = FALSE)+
    labs(x="Binned time", title = "Subject 1 out-of-sample MLE"),

@@

@@

@@

@@

#### Predictive performance measures

- Modeled after dynamic prediction paper (Leroux 2017), focusing on prediction of individual tracks
- Integrated error: for latent process

$$\frac{1}{N}\sum_{i=1}^N\sum_s(\hat{\eta}_{is}-\eta_{is})^2$$
where s covers only unobserved tracks

- Integrated AUC: for binary outcome
- Also, compare with a existing method (e.g, AMM, FRI)
- Perhaps, also computation time?

<span style="color:red;">The problem with measures above is that predictions are made on a different grid than observations. I can think of two potential ways to address this issue: 1) only evaluate at bin midpoints; 2) use interpolation to extend latend function from binned grid to the original grid (in GFPCA paper this is done through projection. ) </span>

```{r}
df_true_eta %>% filter(id==1)
eta_hat_mat <- matrix(NA, ncol = )

out_samp_dyn_pred(df_new = df %>% filter(id==1 & sind_inx<=195) %>%
                                  select(-eta_i),
                                fpca_fit = fpca_fit)$eta_pred
```

@@

Some new updates:

Actually, after going over the full derivation in Leroux2017 paper, I think the preidction interval completely ignored the randomness from fixed effects. In other worlds, prediction intervals are made conditioning on fixed effect coefficients (in our case $\boldsymbol{f_0}$).
